### 2018-02-26 thesis/contents/discussions.tex
Updates from ShareLaTeX

\item Overcoming the major technical barrier to diffusion tractography study of the trigeminal pathway in a large population group, by developing and validating the Selective Automated Group Integrated Tractography (SAGIT) software platform. For the process, we determined the optimal T1-DWI co-registration methodology with clinical DWI sequences. Prior to this project, the optimal co-registration method had not been exhaustively studied.

\item Merged group tractography study of the trigeminal nerve and its primary sensory pathway as a novel use of Gaussian Process machine learning to the study of TN. We studied the trigeminal primary (CN V), secondary (trigeminopontothalamic, TPT), and tertiary (thalamocortical, S1) white patter pathways in sub-regional detail. Our report defines distinct bilateral changes in trigeminal neuralgia subjects, as well as changes between affected and unaffected sides.

@@

\section{Trigeminal Neuralgia}

The findings in Study I (Chapter \ref{section:study1}), which aimed to identify trigeminal nerve diffusivity signatures in TN and MS-TN, identified for the first time a method that can distinguish between microstructural changes in these conditions, and revealed localized CN V diffusivity differences in MS-TN and idiopathic TN. The possible implication is that disruption anywhere along both the cisternal and pontine segments of the trigeminal nerve may lead to similar TN-like pain syndromes. Central demyelination in the trigeminal root transition zone is associated with vascular compression in TN. Moreover MS is a classic myelin inflammatory disease. Therefore both diseases may share pathophysiology due to disruptions in myelin ultrastructure. Prior to the study, identification of the CN V nerve inside the brainstem from MR images were highly inconsistent, and the precision of multiple ROI placements could not be maintained. Multi-tensor tractography was instrumental in the feasibility of the method. The study opened the possibility that multi-tensor tractograpy aided analysis added a measure of objective assessment in examining pain conditions, and may facilitate similar studies that provide insight into the fundamental changes that occur at the level of white matter.

Study IV (Chapter \ref{section:study4}), which applied group tractography to the study of the trigeminal sensory pathway, further suggested that the unaffected CN V diffusivity is also different from that of controls. DeSouza et al. \cite{Desouza2013} also presented similar findings, where the unaffected RD, MD, and AD were found to differ from controls. Desouza performed the study with manual ROI placement at a single measurement position in the REZ, whereas we performed GP classification along the nerve automatically. The GP classifier produced results by exhaustively examine moving windows at all position and sizes, and the results are auto-discovered by the algorithm, in contrast to the prior hand-placement of the measurement ROI in previous studies. The consistency in CN V findings in the location and diffusivity across both manual and automatic methods, suggest that the evidences for diffusivity changes in the unaffected CN V are convincing. This finding, however, is not surprising. It has been long observed that bilateral neurovascular compression could be observed in some unilateral TN cases. Moreover, in third-degree neurovascular compression, severe vascular hardening and deformation can be observed in the cisternal space. It is reasonable that these observable macroscopic disturbances can effect CNS tissue changes on the unaffected side. Moreover descending modulation from PAG can become maladaptive or ineffective in chronic pain, and in these cases more centrally involved changes in the trigeminal nucleus and its associated pathways are possible.
The unaffected side was previously presumed to be similar to normal tissue due to the absence of pain. In many studies it has also been used as a basis of comparison. With this additional insight, it becomes important to not interpret results of affected-to-unaffected comparisons in terms of differences between pathological and healthy tissue. Additionally, the logical next question is to determine that, if the unaffected nerve differs from controls, at what degree of difference does it result in pain. Investigations into bilateral TN pain, and vascular compression in the absence of pain, may reveal additional insight into the relationship between white matter changes and TN pain in general.

@@

The study of trigeminal neuralgia has revolutionarized the approach to the investigation of this unique pain disorder. Additionally, an important motivation for the use of tractography to study of the trigeminal nerve was the ability to overcome the significant limitations of current conventional MR imaging in the visualization of the brainstem trigeminal nerve and the area of the trigeminal nucleus. Prior to the use of tractography, the fibers of the trigeminal nerve and connected pathways could only be inferred from anatomical studies or postmortem dissections.

@@

Diffusion tensor imaging revolutionized how we examine brain white matter fibers, yet resolving cross-fibers of the cerebral white matter is a limitation of diffusion tractography. The brain white matter consists of complex crossing fiber crossings, and the inability for DTI to reliably image these structure therefore posed a fundamental challenge for neuroimaging.
Techniques to improve visualization of white matter fibers consist in probabilistic tractography, which used tissue priors to sample a large number of directions at each step stochastically, and combined the final result as a volumetric probability map. This kind of image-based probabilistic tractography was useful for whole-brain white matter studies. However, it was constrained to the resolution of the diffusion acquisition, which was insufficient for localized and small diameter pathways. Moreover, the algorithm depended on pre-generated tissue-priors, which excluded cranial nerves. 
Deterministic algorithms such as diffusion spectral imaging (DSI), and the related multi-shell acquisition strategies attempted to model the fiber orientation distribution by dramatically increasing the number of gradient directions (256--512 directions were required) in the diffusion acquisition. The acquisition time of these sequences was impractical for the clinical environment. Reduction in acquisition time for these algorithms is currently an active area of research.
The family of spherical harmonics (SH) based methods, which include Q-ball and constrained spherical deconvolution (CSD), were a significant development. SH-based methods modelled the fiber orientation distribution as a decomposition of SH basis functions. This meant that these methods are more robust to noise, and therefore require less number of gradient acquisitions. CSD proved to be particularly robust to noise, and can be deployed on acquisitions with only 55 directions. Another family of tractography (XST) algorithm constrained its model assumption to two fiber crossings, also proved to provide good results with 50 gradient directions.

@@

We divided the trigeminal nerve into 30 discrete cross-sections along its path, and observed detailed diffusivity changes as the nerve transitioned from pontine tissue to the cistern. The number of divisions chosen was arbitrary, with consideration that the number 30 could easily be divided into 3 sub-segments to represent the beginning, middle, and end of an elongated structure. In theory the number of divisions could be infinitely large, given that a centroid curve was identified. In practice the sub-sectional volume decreased with more divisions, and number of points that were sampled by each cross-sectional volume needed to be considered. Determining a common centroid curve from many subjects was computationally non-trivial, especially when the streamline lengths were highly variable due to the noise in the cistern CSF. White matter pathways are highly complex, with possible multi-directional forks, bends, and convergent pathways across a wide range of scales, some examples include the forniceal crura, zona incerta, and optic chiasm. A generalized algorithm that can reliably calculate a centroid for these structure is still a topic of research. 
Turns out that the merged tractography strategy simplified this task by allowed a single curve to be defined manually in the template tractography space. This greatly reduced the amount of labour for researchers, as centroids no longer need to be defined repeatedly for each subject. Manually curve definition by experts can leverage known techniques in computer graphics to produce centroid for very complex structures, thereby extending SAGIT for obtaining along-the-track measure for other highly variable streamline bundles.

The segmentation of the subregions of the trigeminal pathway has allowed for the possibility of measuring the respective diffusivities. Moreover, this method can extend to any scalar images that can be registered to the DWI space. These may include specific tissue-metrics such as myelin, G-ratio, and PET images. This methodology is therefore highly applicable and not specific to the study of TN.

@@

We applied machine learning to the analysis of TN in Study IV in order to avoid our own confirmation bias. The analysis of the trigeminal nerve in TN versus controls served as validation for the study's method. We had performed traditional multiple comparison statistical analysis, but found the result hard to interpret. This was largely due to difficulties in determining the right p-threshold, where the choice of p-value can result in profoundly different statistics. Eventually it was decided that frequentist statistics was not suitable to our dataset, as we are burdened with too great amount of data, which may not fit with frequentist assumptions. The GP classifier was chosen, over more popular SVM classifiers, for the following reasons: 1) the obtained metrics show high location dependency. 2) SVM results are hard to interpret, whereas GP is a bayesian method, and is much more intuitive. More specifically, the streamline measurements resemble time-series data, in that it exhibits positional dependency between measurements. Unlike a time-series however, the streamline measurements represent physical locations along a pathway, therefore they exhibit bi-directional dependency, as suppose to uni-directional in time. Posing the study as a supervised machine learning problem, the feature vector would be a N dimension vector representing N measures at each division, and a classification label. SVM would assume the features to be points in a N-dimensional space, and each measurement would assume independence. This makes the result of the SVM output difficult to interpret. The GP classifier models the feature vector as a latent GP model, before squashing it to produce a logistic output. Therefore the GP model is able to take the positional dependency into account, and the result output is also much more intuitive. Moreover, changing the GP classification model into a regression one is trivial. This importance is illustrated thus: Study IV asked the question of at which measurement positions can we maximally discriminate TN from controls at a group level, but it is also possible to pose a prognosis question: Given N measurements along a patient's CN V, what is the probability that is resembles a pathological population, what about a healthy population? It is possible to answer these questions trivially with the GP model using regression. The full implication is outside of the scope of this thesis, but its implementation is has already been seeded within.

\section{Neuroimaging of deeper white matter for the study of pain}

The findings in S1 in Study IV (\ref{section:study4}) showed TN diffusivity differences in regions surrounding the thalamus, as well as in the mid-segments of the S1 pathways. These findings were in agreement with pain literature, where thalamus and S1/S2 activations are often reported as part of the pain neuromatrix. The TPT pathway resulted in differences near the TGN and the thalamus, areas that are expected to be involved in TN pain. It would be of great interest to re-examine the MS-TN population in these pathways, in order to determine any relationship between deeper white-matter MS-lesion, MS-TN, and non-TN pain in the people with MS.

We have shown in Study III (\ref{section:study3}), for the first time, that different tractography algorithms are differentially suited to delineate different neuropathways when comparing merged reconstructions in 42 subjects. Maier-Hein et al came to similar conclusions after examining 96 reconstructions from a single set of simulated data \cite{Maier-Hein2017}.
Therefore care must be taken to ensure tractography parameters are targeted to the anatomy of study, as more consistent delineation lead to more robust tissue measurements.

Quantification of tissue measures from tractography, as demonstrated in Study IV, implies the use of points in tract models as a sampling method for the scalar values in DWI space. We demonstrated in Study III, that different tratography algorithms exhibit different geometric point distributions. Eventhough we average the point-measures during along-the-tract analysis, which may be considered a form of smoothing/down-sampling. It is not clear whether consistency in comparing measures drawn from the same sampling distribution should be considered as a factor. In other words, given two tracts generatef from identical ROI definitions, with similar cross-sectional area, but of different algorithm, are comparisons between them still valid. Intuitively the algorithm choices and parameters should be identical, and that's what we did in Study III. However, there maybe cases where different parameters are better suited for certain pathologies, and this is may be a concern when deploying tractogrpahy analysis towards clinical diagnosis in the future, and therefore merit further investigation.

\section{Neurosurgery and Tractography}

\section{Method Discussions}

@@

From my observations across the studies in TN, I postulate a resonance hypothesis of TN. I am aware that there lacks substantial physiological measurements for these claims, nevertheless I feel it is important to state them here as a possible direction of investigation for the future: In surgical interventions for TN, the target is on the cistern CN V segment, and situates close to the CN V REZ. Paradoxically, the intervention can be seen as a willful disruption of the CN V myelination. The question is, why would additional disruptions on the CN V myelin effectively decrease TN pain. Ephaptic transmission of action potential from the fast conducting fibers to the slow conducting pain fibers is thought to explain TN pain. The paroxysmal nature of TN pain, and the unreliable association between compression and symptom, however, suggest additional factors are involved. Consider that blood circulation and motion is constant within the cistern space, the trigeminal-vascular complex is a dynamically undulating structure. The force exerted onto the CN V from its neighbouring vasculature is therefore periodic. It is possible that during the period where the vascular pressure onto the CN V is at peak, and at the same time the action potential conductance is near the locus of compression, that the ephaptic transmission is triggered and amplified, creating a "resonance condition". This would imply that given periodic vascular forces, and a constant stream of innocuous afferents, the chance of resonance condition is small, and can result in no observable pattern or interval, thus explaining the paroxysmal nature of TN pain. Moreover, under bilateral vascular compression, only the side with vascular and CN V physiology that can result in elevated chance of such resonance condition will result in pain. This model also explains that surgical intervention disrupts the chance of resonance condition, thereby reducing TN pain. The pain eventually can return however, as the peripheral lesion site heals and the resonance condition returns. MVD, in this case can be explained as the physical disruption of the condition.

>>>>

\item Overcoming the major technical barriers to diffusion tractography study of the trigeminal pathway in a large population group, by developing and validating the Selective Automated Group Integrated Tractography (SAGIT) software platform. For the process, we determined the optimal T1-DWI co-registration methodology with clinical DWI sequences.

\item Merged group tractography study of the trigeminal nerve and its primary sensory pathway as a novel use of Gaussian Process machine learning to the study of TN. This study involved spatial defintion of groups tracts in space, as well as measurement of diffusivity metrics from the subregions of trigeminal primary (CN V), secondary (trigeminopontothalamic, TPT), and tertiary (thalamocortical, S1). We demonstrate distinct bilateral changes in trigeminal neuralgia subjects, as well as differences between affected and unaffected sides.

@@

\section{A new approach to the study of Trigeminal Neuralgia}

The introduction of DTI in the study of TN has revolutionized our understanding of this disorder. Moving away from clinical studies alone, DTI is able to derive metrics that provide important information on the microstructure of the nerve which can then be related to specific processes such as changes in myelination or the microstructure of the axons. Thus DTI provides an important objective measure in the study of pain, which was unavailable up unto this point. The earliest work on the feasibility of accurate reconstruction of cranial nerve with DTI tractography \cite{Hodaie2010} in 2010. This study was able to image CN II, III, and importantly CN V. Similarly, we were able to establish the methods to image the TRG and the cisternal CN V with manual delineation, followed by visualization of the CN V postganglionic GK target, prior and after surgery \cite{Hodaie2012g}. We found alternations in both the CN V reconstruction, as well as FA in the surgical target location, that improved our understanding beyond what is visible with post-surgical gadolinium enhancements. This has now led our team to study whether CN V GK target diffusivity at 6 months post-surgery can predict long-term pain relief in TN \cite{Tohyama2018}. Similarly whether the diffusivity changes are specific enough that they can serve as predictors of outcome \cite{Hung2017}. The tractography study of TN has been directly linked with the study of gray and white matter abnormalities of the TN brain \cite{Desouza2013c}. The value of DTI studies in TN is therefore unique, and creates a strong context for the work done in this thesis.

The findings in Study I (Chapter \ref{section:study1}), point to the clear role of DTI in distinguishing the localized, focal changes in diffusivity that characterize TN and MS-TN. While TN is associated with compression of the REZ and possible associated focal central demyelination \cite{Devor2002a,Peker2006}, MS is associated with inflammation and myelin changes resulting in CNS plaques. While both diseases involve the same clinical expression of pain, the pathophysiology of pain differs, but this had not been shown directly using DTI based imaging techniques. From a technical perspective, the identification of the CN V nerve inside the brainstem from MR images had been previously highly inconsistent, and the precision of multiple ROI placements could not be maintained. Multi-tensor tractography was instrumental in the feasibility of the method. The study opened the possibility that multi-tensor tractograpy aided analysis added a measure of objective assessment in examining pain conditions, and may facilitate similar studies that provide insight into the fundamental changes that occur at the level of white matter.

As a result of the methodology developed in Study I, a recent publication details the investigation of whether presurgical CN V segment diffusivity values can predict response to TN surgery \cite{Hung2017}. The important study found that lowered cistern MD and AD pre-surgery was correlated with positive surgical outcome, whereas lowered REZ FA and increased pontine AD was correlated with the lack of response from surgery. This provided further evidence of the role of cisternal CN V diffusivities in the study of TN. The fact that the novel finding of pontine diffusivity in classic TN favours a non-responder outcome points to the role of CNS alterations in classic TN.

Study IV (Chapter \ref{section:study4}), applied group tractography to the study of the trigeminal sensory pathway, and highlighted the observation that the unaffected CN V diffusivity in TN differs from that of controls. There have been other studies suggesting bilateral abnormalities in TN. A group level diffusivity study of TN was conducted by DeSouza et al. in 2013 \cite{Desouza2013} presented similar findings. DeSouza manually defined ROI in the trigeminal REZ, as well as performing full-brain Tract-Based Spatial Statistics. Bilateral differences in the REZ between TN patients and controls were found in MD, RD, and AD. Wide-spread white matter differences were also found, especially in the corpus callosum, cingulum, and longitudinal fasciculus. DeSouza had also found increased cortical and grey-matter volume changes in TN patients \cite{Desouza2013c}, including increase in contralateral S1 and thalamus volumes.
Using SAGIT in study IV, the ROIs were projected to each individual subject using an automated pipeline. The GP classifier produced results by exhaustively examine moving windows at all position and sizes, and the results are auto-discovered by the algorithm, in contrast to manual-placement of the measurement ROI in previous studies. There is therefore consistency in CN V findings in the location and diffusivity across both manual and automated methods.

The observation of bilateral differences has important implications in TN pain. From previous studies, it is clear that altered diffusivities in the cisternal and pontine segments of CN V are correlated with TN pain in the affected side. However there are now a number of studies that point to bilateral REZ/pontine CN V diffusivities being altered in unilateral TN, although each point to significant differences between affected and unaffected sides as well. The exact nature of bilateral diffusivity alterations is not yet clear. Perhaps the earliest indication that TN can affect the nerves bilaterally is a clinical paper by the Burchiel group, where they point to a surprisingly large number of patients with unilateral TN that have evidence of bilateral neurovascular compression, in contrast with healthy controls \cite{Miller2009}. Moreover descending modulation from PAG can become maladaptive or ineffective in chronic pain, result in central changes in the trigeminal nucleus and its associated pathways.

The unaffected side was previously presumed to be similar to normal tissue due to the absence of pain. In many studies it has also been used as a basis of comparison. With this additional insight, it becomes important to not interpret results of affected-to-unaffected comparisons in terms of differences between pathological and healthy tissue. Additionally, the logical next question is to determine that, if the unaffected nerve differs from controls, at what point does the clinical expression of pain start. Investigations into bilateral TN pain, and vascular compression in the absence of pain, may reveal additional insight into the relationship between white matter changes and TN pain in general. The newly developed SAGIT and machine learning methods can additionally be leveraged to conduct large cohort studies, that allow investigation into the brain regions identified by DeSouza et al, as well as examining important areas in the modulation of pain, such as the insula and its relationship to TN.

Finally, with the methods developed throughout this thesis, future studies are no longer restricted by manual ROI placements for each subject, and tract subregions can be studied in detail. This opens the avenue to the study of associated pathways of TN pain, and allow the development of more sophisticated prediction models to further the understanding, and treatment of TN.

@@

An important motivation for the use of tractography to study the trigeminal nerve was the ability to overcome the significant limitations of current conventional MR imaging in the visualization of the brainstem trigeminal nerve and the area of the trigeminal nucleus. Prior to the use of tractography, the fibers of the trigeminal nerve and connected pathways could only be inferred from anatomical studies or postmortem dissections.

@@

Diffusion tensor imaging revolutionized how we examine brain white matter fibers, yet resolving cross-fibers of the cerebral white matter remains a limitation of diffusion tractography. The brain white matter consists of complex crossing fiber crossings, and the inability for DTI to reliably image these structure therefore poses a fundamental challenge for neuroimaging.
Techniques to improve visualization of white matter fibers consist of probabilistic tractography, which uses tissue priors to sample a large number of directions at each step stochastically, and combines the final result as a volumetric probability map. This kind of image-based probabilistic tractography has proved useful for whole-brain white matter studies. However, it is constrained to the resolution of the diffusion acquisition, which can be insufficient for localized and small diameter pathways. Moreover, the algorithm depends on pre-generated tissue-priors, which excludes cranial nerves. 
Deterministic algorithms such as diffusion spectral imaging (DSI), and the related multi-shell acquisition strategies attempt to model the fiber orientation distribution by dramatically increasing the number of gradient directions (256--512 directions where required) in the diffusion acquisition. The acquisition time of these sequences can be impractical for the clinical environment. Reduction in acquisition time for these algorithms is currently an active area of research.
The family of spherical harmonics (SH) based methods, which include Q-ball and constrained spherical deconvolution (CSD), are a significant development. SH-based methods model the fiber orientation distribution as a decomposition of SH basis functions. This means that these methods are more robust to noise, and therefore require less number of gradient acquisitions. CSD proved to be particularly robust to noise, and can be deployed on acquisitions with only 55 directions. Another family of tractography (XST) algorithm constrained its model assumption to two fiber crossings, also proved to provide good results with 50 gradient directions.

@@

As part of our focusing on ??? and for the purposes of imaging the trigeminal pathways, where a significant number of crossing fibers exist, we divided the trigeminal nerve into 30 discrete cross-sections along its path, and observed detailed diffusivity changes as the nerve transitioned from pontine tissue to the cistern. The number of divisions chosen was arbitrary, with consideration that the number 30 could easily be divided into 3 sub-segments to represent the beginning, middle, and end of an elongated structure. In theory the number of divisions could be infinitely large, given that a centroid curve was identified. Determining a common centroid curve from many subjects was computationally non-trivial, especially when the streamline lengths are variable. White matter pathways are highly complex, with possible multi-directional forks, bends, and convergent pathways across a wide range of scales. A generalized algorithm that can reliably calculate a centroid for these structure is of great value and yet still a topic of research. 
Merged tractography  simplified this task by allowing a single curve to be defined manually in the template tractography space. This greatly simplied the task, since centroids no longer need to be defined repeatedly for each subject. Manually curve definition by experts can leverage known techniques in computer graphics to produce centroid for very complex structures, thereby extending SAGIT for obtaining along-the-track measure for other highly variable streamline bundles.

The segmentation of the subregions of the trigeminal pathway has allowed for the possibility of measuring the respective diffusivities. Moreover, this method can extend to any scalar images that can be registered to the DWI space. These may include specific tissue-metrics such as myelin and G-ratio images. This methodology is therefore highly applicable and not specific to the study of TN.

\section{Neuroimaging of deep white matter}

The findings in S1 in Study IV (Chapter \ref{section:study4}) showed TN diffusivity differences in regions surrounding the thalamus, as well as in the mid-segments of the S1 pathways. These findings were in agreement with pain literature, where thalamus and S1/S2 activations are often reported as part of the pain neuromatrix. The TPT pathway resulted in differences near the TGN and the thalamus, areas that are expected to be involved in TN pain. It would be of great interest to re-examine the MS-TN population in these pathways, in order to determine any relationship between deeper white-matter MS-lesion, MS-TN, and non-TN pain in the people with MS.

We have shown in Study III (Chapter \ref{section:study3}) that different tractography algorithms are suited to delineate different neuropathways when comparing merged reconstructions in 42 subjects. Maier-Hein et al came to similar conclusions after examining 96 reconstructions from a single set of simulated data \cite{Maier-Hein2017}. As tractography is applied to the investigation of white matter elsewhere in the brain, care must be taken to ensure  parameters are targeted to the anatomy of study, as more consistent delineation leads to more robust tissue measurements.

Quantification of tissue measures from tractography, as demonstrated in Study IV, implies the use of points in tract models as a sampling method for the scalar values in DWI space. We demonstrated in Study III, that different tractography algorithms exhibit different geometric point distributions. In other words, given two tracts generated from identical ROI definitions, with similar cross-sectional area, but of different algorithm, the underlying spacial distribution of the points are different. It can be argued that we average the point-metrics during along-the-tract analysis, which may be considered a form of smoothing/down-sampling, and therefore the spatial distribution of the tractography algorithm might not be important. However, until the exact influence of spatial distribution have on the measured distribution of diffusivity metrics can be understood, comparing between similar tracts from two different tractography algorithms should be avoided.

@@

The availability of the SAGIT, full automation of tractography processing enabled a dramatic increase in the volume of data available for statistical analysis. This lead to some challenges in the scalability of the analytical process, but also opportunities in leveraging more expressive machine learning approaches. The need for data to improve diagnosis and treat of disease, particularly in the emergence of artificial intelligence in medical imaging is increasingly been recognized. Data is now the foundation from which new techniques and discoveries are possible. Initiatives such as the UK Biobank, Allen Brain Institute, and Ontario Brain Institute all collect and provide their neuroimaging data to researchers. With this deluge of imaging data, the complexity of analysis that demand of researchers proportionally increases. The need to include large amount of variables, patient attributes, study measures, time sequences, can now provide valuable insight that is beyond the mean of traditional statistics. In this thesis we developed an machine learning approach that is scalable, and one that can quantify uncertainty as demanded by the medical need. The data processing and machine learning is end-to-end, meaning that the system can be practically deployed in a clinical setting.

We applied machine learning to the analysis of TN in Study IV. The analysis of the trigeminal nerve in TN versus controls agreed with our earlier findings, and therefore validates the study's GP classification method. We performed traditional multiple comparison statistical analysis for our model, but found the result hard to interpret. This was largely due to difficulties in determining the right p-threshold, where the choice of p-value can result in profoundly different statistics. We therefore decided that frequentist statistics was not suitable to our dataset, as we are burdened with too great amount of data, which may not fit with frequentist assumptions. 
The GP classifier was chosen over more popular SVM classifiers, for the following reasons: 1) the obtained metrics show high location dependency. 2) SVM results are hard to interpret, whereas GP is a bayesian method, and therefore it is much more intuitive. More specifically, the streamline measurements resemble time-series data, in that it exhibits positional dependency between measurements. Unlike a time-series however, the streamline measurements represent physical locations along a pathway, therefore they exhibit bi-directional dependency, as opposed to uni-directionally dependent in time. Posing the study as a supervised machine learning problem, the feature vector would be a N dimension vector representing N measures at each division, and a classification label. SVM would assume the features to be points in a N-dimensional space, and each measurement would assume independence. This makes the result of the SVM output difficult to interpret. The GP classifier models the feature vector as a latent GP model, before squashing it to produce a logistic output \cite{rasmussen2006gaussian}. Therefore the GP model is able to take the positional dependency into account. 
Changing the GP classification model into a regression one is very simple. This importance is illustrated thus: Study IV asked the question of at which measurement positions can we maximally discriminate TN from controls at a group level, but it is also possible to pose a prognosis question: Given N measurements along a single patient's CN V, what is the probability that is resembles a pathological population as opposed to a healthy population? It is possible to answer these questions trivially with the GP model using regression. The full implication is outside of the scope of this thesis, but this work will serve as a strong baseline for the applications in this direction.

\section{Novel Technical developments}

@@
